\textbf{\sc Towards Deliberative Control of Autonomous Robots}

### 2012-02-21 trunk/europa-basics.tex
europa-basics cleanup.

% . In
% this section we detail some of the key concepts relevant to flexible
% plan representation. Section \ref{sec:arch} has details the \rx
% partitioned architecture and estimation driven vehicle adaptation.

% \begin{figure}
% \centering
% \includegraphics[scale=0.3]{figs/Allen-algebra.pdf}
% \caption{\small Temporal relations defined within the planner are
%   based on \texttt{Allen Algebra} relations shown above.}
% \label{fig:allen-algebra}
% \vskip-0.3cm
% \end{figure}

\subsection{Fundamentals of  Automated Planning}
\label{sec:planningfound}

@@

% Planning, for our purposes, can be thought of as determining all the
% small tasks that must be carried out in order to accomplish a goal.
To articulate the fundamentals of automated planning briefly and use
that to motivate the mechanisms we use in our specific form of the
technique we start with a simple example.

@@

\begin{eqnarray*}
 Sum(Q_{rc}) & = & \left\{
   \begin{array}{l l}
     1 & \forall r \quad \mbox{(only one Queen per row)}\\
     1 & \forall c \quad \mbox{(only one Queen per column)}\\ 
   \end{array} \right. \\
Sum(Q_{r+i,c+i}) & = & 1\\
Sum(Q_{r-i,c+i}) & = & 1 \qquad \mbox{(only one Queen on each diagonal)}\\
\end{eqnarray*}

%\comment{fix newline and numbering problem above}

The problem can be solved by finding assignments for all variables
Q$_{rc}$ that satisfy the above constraints. Fig. \ref{fig:nqueens-2} is
a solution found by \eu using that formulation and a specialized search
procedure.

@@

Constraint Satisfaction Programming (also known as Constraint
Programming (\textsf{CP})) is a discipline that provides a generic framework
for representing, solving and making logical inference on
constraints. A complete treatment of this discipline is given in
\cite{marriott98,apt03} and very concise introductions are provided in
\cite{bartak99,lustig01}.

@@

\begin{figure}[!t]
\centering
\includegraphics[scale=0.35]{figs/constraint-net.pdf}
\caption{\small Illustration of a constraint network as represented as
a graph.}
\label{fig:constraintnet}
\end{figure}

A set of constraints and variables that describe a problem domain are
typically represented as a Constraint Network, where the variables are
nodes and constraints are arcs between the variables. Having this
network representation in mind, a pair of variables $x,y$ is said to
be arc consistent if for every value in $x$'s domain, there is a value
in $y$'s domain that is consistent with all the constraints that
connect $x$ and $y$.  A naive approach to ensuring arc consistency
consists of cycling through all the variable pairs and performing
constraint propagation until there are no variable domain changes. The
\texttt{AC-3} algorithm \cite{mackworth77}, a popular implementation,
improves over the naive approach by ignoring constraints whose
variables were not affected since the last
iteration. Fig. \ref{fig:constraintnet} shows an example of a
constraint network with three variables, their domains and the
constraint relations between them.

@@

The most common planning formulations use a propositional
representation, where the state of the world is represented by a set
of propositions (statements that can be true or false), and operators
change the truth values of these propositions \cite{gen87}. Although
these formulations are powerful and have allowed researchers to
develop numerous contributions in automated planning, there are many
classes of problem domains that are difficult to represent using this
formalism. In particular, it is hard to represent time, resources,
% mutual exclusion and concurrency using propositions
% \cite{ghallab04}. While CSP representations have traditionally had an
% edge in formulating and representing planning problems,
% % It is straightforward to represent and reason about all of
% % those elements using a CSP representation, as a result there has been
% % some work on doing automated planning while taking advantage of CSP
% % (TODO: ref). Traditionally, the entire planning problem is translated
% % into a CSP and then solved using traditional CSP methods, this
% leading to a formulation where action choices and relationships are
% represented as variables and/or constraints, there are at least two
% major drawbacks:
mutual exclusion and concurrency using propositions \cite{ghallab04}.
Constraint Programming is a framework that lends itself very well to
represent and reason about all of those elements \cite{ghallab94},
consequently, there has been a good amount of work building automated
planners that take advantage of CP techniques.  The most common
approach in constraint-based planning has been to encode the entire
planning problem as a CP problem, where action choices and
relationships are represented as variables and/or constraints
\cite{do01,vanbeek99,vossen99,wolfman99}.  The advantage of using a
pure CP formulation is that standard CP solvers can be then used to
solve the planning problem, however, this approach also has a couple
of significant drawbacks:

@@

\item If the structure of a planning model (actions, conditions,
  effects, dependencies at the action level) is not explicitly
  maintained by the CSP planner, the search algorithms are deprived of
  critical information to make better decisions. If that structure is
  maintained (for instance, by internally marking variables that
  represent action choice and relationships between actions), it is
  still hard to write search algorithms as any planning-specific
  insight has to be translated into the CSP representation of
  variables and constraints \comment{For example?}

@@

These drawbacks have been addressed by a small number of planning
frameworks like \texttt{Descartes} \cite{joslin95,joslin96},
\texttt{IxTet} \cite{ghallab87,lemai04}, and CAIP (Constraint-Based
Interval and Attribute Planning)\cite{mus94,frank2003}.  These
frameworks take advantage of a CSP representation to deal with time
and resources, at the same time maintaining an explicit representation
of the elements relevant to planning (action choices, causal
relationships, etc). This combination results in a planning approach
that can deal with the type of constraints found in real-life
problems, while allowing for intuitive problem descriptions and a
natural way to express search algorithms and heuristics. \eu is an
implementation of the CAIP framework. In CAIP, planning problems are
specified in terms of:

@@

To generate plans, the CAIP framework uses a plan-space search
approach \cite{ghallab04}, where a partial plan is evolved until all
open goals are achieved, and all inconsistencies are removed, the
specific search algorithm used by \eu will be covered in detail below.

>>>>

The next few sections will cover \eu in a bottom-up fashion, going from the 
generic theoretical frameworks that underlie \eu, to the specific representation, 
modeling and algorithms implemented by it. First, a quick introduction to 
Constraint Programming and Constraint-Based Planning is provided. 
Next, the specific representation that \eu uses to perform Constraint-Based Planning is presented.
After that comes an overview of \eu's architecture and how it can be embedded in a planning application
such as \rx. Finally,  \eu's modeling, inference and search mechanisms are covered in 
detail so that the reader can understand how \rx can perform deliberative planning and react to
execution events by embedding and customizing \eu.

\subsection{Automated Planning and \eu}
\label{sec:planningfound}

@@

To articulate the fundamentals of automated planning briefly and use
that to motivate the mechanisms we use in our specific form of the
technique we start with a simple example.

@@

\begin{eqnarray*}
 Sum(Q_{rc}) & = & \left\{
   \begin{array}{l l}
     1 & \forall r \quad \mbox{(only one Queen per row)}\\
     1 & \forall c \quad \mbox{(only one Queen per column)}\\ 
   \end{array} \right. \\
Sum(Q_{r+i,c+i}) & = & 1\\
Sum(Q_{r-i,c-i}) & = & 1 \qquad \mbox{(only one Queen on each diagonal)}\\
\end{eqnarray*}

The problem can be solved by finding assignments for all variables
Q$_{rc}$ that satisfy the above constraints. Fig. \ref{fig:nqueens-2} is
a solution found by \eu using a compact version of that formulation and a specialized search
procedure.

@@

Constraint Satisfaction Programming (also known as Constraint
Programming or \textsf{CP}) is a discipline that provides a generic framework
for representing, solving and making logical inference on
constraints. A complete treatment of this discipline is given in
\cite{marriott98,apt03} and very concise introductions are provided in
\cite{bartak99,lustig01}.

@@

A set of constraints and variables that describe a problem domain are
typically represented as a Constraint Network, where the variables are
nodes and constraints are arcs between the variables. Having this
network representation in mind, a pair of variables $x,y$ is said to
be arc consistent if for every value in $x$'s domain, there is a value
in $y$'s domain that is consistent with all the constraints that
connect $x$ and $y$.  A naive approach to ensuring arc consistency
consists of cycling through all the variable pairs and performing
constraint propagation until there are no variable domain changes. The
\texttt{AC-3} algorithm \cite{mackworth77}, a popular implementation, improves
over the naive approach by ignoring constraints whose variables were
not affected since the last iteration. Fig.  \ref{fig:constraintnet} shows a set of constraints on 3 variables
($x$, $y$ and $z$) represented as a constraint network, and their domains before arc consistency is enforced.
After the AC-3 algorithm is executed, the domains for the 3 variables
would be pruned to x=$\{2$\}, y=$\{2$\}, z=$\{2$\}.

\begin{figure}[!t]
\centering
\includegraphics[scale=0.35]{figs/constraint-net.pdf}
\caption{\small Variables and Constraints represented as a Constraint Network}
\label{fig:constraintnet}
\end{figure}

@@

The most common planning formulations use a propositional
representation, where the state of the world is represented by a set
of propositions (statements that can be true or false), and operators
change the truth values of these propositions \cite{gen87}. Although
these formulations are powerful and have allowed researchers to
develop numerous contributions in automated planning, there are many
classes of problem domains that are difficult to represent using this
formalism. In particular, it is hard to represent time, resources,
mutual exclusion and concurrency using propositions \cite{ghallab04}.
Constraint Programming is a framework that lends itself very well to
represent and reason about all of those elements \cite{ghallab94},
consequently, there has been a good amount of work building automated
planners that take advantage of CP techniques.  The most common
approach in constraint-based planning has been to encode the entire
planning problem as a CP problem, where action choices and
relationships are represented as variables and/or constraints
\cite{do01,vanbeek99,vossen99,wolfman99}.  The advantage of using a
pure CP formulation is that standard CP solvers can be then used to
solve the planning problem, however, this approach also has a couple
of significant drawbacks:

@@

\item If the structure of a planning model (actions, conditions,
  effects, dependencies at the action level) is not explicitly
  maintained by the CSP planner, the search algorithms are deprived of
  critical information to make better decisions. If that structure is
  maintained (for instance, by internally marking variables that
  represent action choice and relationships between actions), it is
  still hard to write search algorithms as any planning-specific
  insight has to be translated into the CSP representation of
  variables and constraints.

@@

These drawbacks have been addressed by a small number of planning
frameworks like \texttt{Descartes} \cite{joslin95,joslin96},
\texttt{IxTet} \cite{ghallab87,lemai04}, and \texttt{CAIP} (Constraint-Based
Interval and Attribute Planning)\cite{mus94,frank2003}.  These
frameworks take advantage of a CSP representation to deal with time
and resources, at the same time maintaining an explicit representation
of the elements relevant to planning (action choices, causal
relationships, etc). This combination results in a planning approach
that can deal with the type of constraints found in real-life
problems, while allowing for intuitive problem descriptions and a
natural way to express search algorithms and heuristics. \eu is an
implementation of the \texttt{CAIP} framework. In \texttt{CAIP}, planning problems are
specified in terms of:

@@

To generate plans, the \texttt{CAIP} framework uses a plan-space search
approach \cite{ghallab04}, where a partial plan is evolved until all
open goals are achieved, and all inconsistencies are removed. The
specific search algorithm used by \eu will be covered in detail below.
